Constraining the Interiors of Asteroids Through Close Encounters

### 2022-04-28 paper/appendices.tex
Pulled speed contour plots

% \section{The cadence cut-off}
% \label{app:cadence-tests}

% In section \ref{sec:scan-cadence}, we noted that posterior uncertainty as a function of observation cadence $\Delta t$ appears to increase suddenly near $\Delta t \sim T_\text{cad}=30-40$ min. We mentioned that this cadence cut-off is likely a function both of the rotational period of the asteroid and the time spent near perigee. We study this cut-off in closer detail in this appendix.

% We measure the dependence of the cadence cut-off on rotational period by varying the period $P_\omega$ and reproducing figure \ref{fig:scan-cadence}; i.e., we run many simulations of the changed $P_\omega$ with different cadences and plot the posterior fit uncertainty $\sigma$ as a function of cadence $\Delta t$. Specifically, we use $P_\omega = 5$ hr and 20 hr, roughly double and half of the reference $P_\omega=9$ hr. The resulting curves of $\sigma(\Delta t)$ are shown in figure \ref{fig:cad-period}, with each normalized so that their maximum values is 1 so that changes to $\sigma$ induced by the change in $P_\omega$ that are not a function of $\Delta t$ (i.e., those studied in section \ref{sec:scan-period}), are removed; they are irrelevant here.

% \begin{figure}
%   \centering
%   \includegraphics[height=0.89\textheight]{figs/cad-period.pdf}
%   \caption{1$\sigma$ confidence intervals plotted as a function of cadence $\Delta t$ for periods of 9 hr (the reference asteroid, solid coloured lines), and 5 hours (dotted lines), and 20 hours (dashed lines). Confidence intervals have been normalized for the sake of comparison, so that the maximum for each curve is 1.}
%   \label{fig:cad-period}
% \end{figure}

% We also wish to capture dependence on the time spent by the asteroid near perigee, but we do not wish to change the orbit shape or central body, because this may induce other effects that could also affect the cadence plot. Instead, we (un-physically) manipulate the speed at which the asteroid moves through the orbit. In one instance, the asteroid's path is artificially slowed by a factor of 2, which we denote as $t_\text{spin}/t_\text{orbit}=0.5$; in the other instance, $t_\text{spin}/t_\text{orbit}=2$, meaning that the orbit is sped up. This approach of changing the speed of the orbit is meant to approximate changes to $T_p$ defined in equation \ref{eqn:tp}. The posterior uncertainties for these two cases and the original case are shown as a function of $\Delta t$ in figure \ref{fig:cad-speed}. Again, $\sigma$ is normalized so that $\Delta t$-independent variations are removed. The $t_\text{spin}/t_\text{orbit}=2$ uncertainties begin to fill the prior distribution for large $\Delta t$ for some parameters, which can be seen by discontinuous variation in the dotted line of figure \ref{fig:cad-speed}.

% \begin{figure}
%   \centering
%   \includegraphics[height=0.89\textheight]{figs/cad-speed.pdf}
%   \caption{1$\sigma$ confidence intervals plotted as a function of the orbit time scaling $t_\text{spin}/t_\text{orbit}$ for no scaling (the reference asteroid, solid coloured lines), time doubling (dotted lines), and halving (dashed lines). See text for the definition of $t_\text{spin}/t_\text{orbit}$. Confidence intervals have been normalized for the sake of comparison, so that the maximum for each curve is 1.}
%   \label{fig:cad-speed}
% \end{figure}

% Many of the curves in figures \ref{fig:cad-period} and \ref{fig:cad-speed} show a cadence cut-off at $T_\text{cad} \sim 30-60$ min. Those that show no cut-off (i.e., they merely smoothly increase over $\Delta t$) may in fact exhibit a cut-off $T_\text{cad} > 60$ min, which is off the plot. By comparing $\sigma(\Delta t)$ for a specific parameter, we see that the dashed lines often (but not always) reach the cadence cut-off at lower $\Delta t$ than the solid line. For example, in figure \ref{fig:cad-period}, this effect is especially clear for $K_{22}$, $K_{33}$, $\Im K_{22}$, and $\Re K_{31}$, but the opposite appears for $\Re K_{32}$. The dashed curves are significant in both figures because they represent larger $\sigma$ cases ($P_\omega=5$ hours, or $t_\text{spin}/t_\text{orbit}=2$); section \ref{sec:scan-period} reveals that fast rotators lead to larger $\sigma$, and $t_\text{spin}/t_\text{orbit} > 1$ leads to larger $\sigma$ because of the presence of less perigee data. It therefore appears that small $P_\omega$ and small times spent at perigee move the cadence cut-off to smaller $T_\text{cad}$.

% The dependence of the exact location of the cadence cut-off is likely more complicated than the above analysis indicates. However, large uncertainties induced by small $P_\omega$ or large $t_\text{spin}/t_\text{orbit}$ makes it difficult to produce data which are not affected by the finite size of the prior (as figure \ref{fig:scan-vex} is) to study these regimes. Computation time and random variation in $\sigma$ also limits us. We therefore do not attempt to explore the dependence between $T_\text{cad}$ and the encounter parameters further. If a future analysis does require a detailed understanding of how cadence affects uncertainty for a specific encounter, our simulation and fit process can be run given the parameters of the encounter in question in a similar manner to section \ref{sec:scan-cadence} to yield more information.

@@

In section \ref{sec:general-density}, we mention that the shape of the asteroid can be computed from observations of the density moments and $a_\mathcal{A}$ if a uniform density distribution is assumed. We did not further describe the model in the main text because it is unlike the other three models in the nature of its assumptions, and in the fact that it is non-linear. However, it may be a useful technique so it is mentioned here; we call it the ``surface'' model. A similar model was also studied in Ref.~\cite{BAXANSKY2007756}.

>>>>

\section{The cadence cut-off}
\label{app:cadence-tests}

In section \ref{sec:scan-cadence}, we noted that posterior uncertainty as a function of observation cadence $\Delta t$ appears to increase suddenly near $\Delta t \sim T_\text{cad}=30-40$ min. In the left panel of figure \ref{fig:cad-contour}, we display contour plots of posterior uncertainty $\sigma(K_{\ell m})$ of the fit parameters as a function of both cadence $\Delta t$ and $P_\omega$ (left panel) or the relative orbit speed $t_\text{spin} \ t_\text{orbit}$ (right panel). Both panels show the sudden increase in posterior uncertainty we named $T_\text{cad}$. Now $T_\text{cad}$ is visible as a function of frequency and relative orbit speed.

\begin{figure*}
  \centering
  \includegraphics[width=0.49\textwidth]{figs/cad-period.pdf}\hfill
  \includegraphics[width=0.49\textwidth]{figs/cad-speed.pdf}
  \caption{Contour plots showing posterior uncertainties in the first- and second-order fit parameters as a function of cadence $\Delta t$ and dynamical time scales for the encounter: rotational period $P_\omega$ (\textit{left}) and the relative speed of the orbit (\textit{right}; see text for a definition).}
  \label{fig:cad-contour}
\end{figure*}

The relative orbit speed $t_\text{spin} \ t_\text{orbit}$ was determined by un-physically increasing or decreasing the time at which the asteroid moved through the orbit determined by the equations of motion (but leaving the orbit shape unchanged). The equations of motion affecting the orientation and spin of the asteroid however were unaffected. With this unphysical process, we isolate the effect of the amount of time spent near perigee on posterior uncertainty, without inheriting additional affects that would have been caused by the orbit changing shape.

Figure \ref{fig:cad-contour} demonstrates that ...

@@

In section \ref{sec:density-distro}, we mention that the shape of the asteroid can be computed from observations of the density moments and $a_\mathcal{A}$ if a uniform density distribution is assumed. We did not further describe the model in the main text because it is unlike the other three models in the nature of its assumptions, and in the fact that it is non-linear. However, it may be a useful technique so it is mentioned here; we call it the ``surface'' model. A similar model was also studied in Ref.~\cite{BAXANSKY2007756}.

### 2022-04-28 paper/main.tex
Pulled speed contour plots

Over the past twenty years, the increase in quantity and quality of sensitive all-sky surveys has prompted the discovery of numerous asteroids. Such advances have been made via ground-based surveys such as the Catalina Sky Survey \cite{larson1998catalina}, Pan-STARRS \cite{kaiser2002pan}, and the Lincoln Near-Earth Asteroid Research project (LINEAR) \cite{stokes2000lincoln}, as well as space-based instruments such as the Wide-field Infrared Survey Explorer (WISE) mission \cite{wright2010wide}. Many of these asteroids are relatively small, but some are kilometre-sized and a few are predicted to closely encounter Earth or other planets in the near future. More encounter candidates are likely to be discovered by new efforts such as the Large-aperture Synoptic Survey Telescope (LSST) \cite{tyson2002large}. Their encounters can then be monitored by global ground-based networks such as the Las Cumbres Observatory (LCO) \cite{brown2013cumbres}. Such ground-based monitoring is typically used to derive the rotation period of an asteroid and its surface properties \cite{}\jtd{Ask for reference from Artem}.

@@

\begin{figure*}
  \centering
  \includegraphics[width=0.3\textwidth]{figs/asym-ell-likelihood.pdf}\hfill
  \includegraphics[width=0.3\textwidth]{figs/asym-ell-harmonic.pdf}\hfill
  \includegraphics[width=0.3\textwidth]{figs/asym-ell-lumpy.pdf}

  % \includegraphics[width=0.3\textwidth]{figs/asym-ell-diff-likelihood.pdf}\hfill
  % \includegraphics[width=0.3\textwidth]{figs/asym-ell-diff-harmonic.pdf}\hfill\hspace{0.3\textwidth}

  \includegraphics[width=0.3\textwidth]{figs/asym-ell-unc-likelihood.pdf}\hfill
  \includegraphics[width=0.3\textwidth]{figs/asym-ell-unc-harmonic.pdf}\hfill\hspace{0.3\textwidth}

  % \includegraphics[width=0.3\textwidth]{figs/asym-ell-ratio-likelihood.pdf}\hfill
  % \includegraphics[width=0.3\textwidth]{figs/asym-ell-ratio-harmonic.pdf}\hfill\hspace{0.3\textwidth}

@@

\begin{figure*}
  \centering
  \includegraphics[width=0.3\textwidth]{figs/blob-likelihood.pdf}\hfill
  \includegraphics[width=0.3\textwidth]{figs/blob-harmonic.pdf}\hfill
  \includegraphics[width=0.3\textwidth]{figs/blob-lumpy.pdf}

  % \includegraphics[width=0.3\textwidth]{figs/blob-diff-likelihood.pdf}\hfill
  % \includegraphics[width=0.3\textwidth]{figs/blob-diff-harmonic.pdf}\hfill
  %\includegraphics[width=0.3\textwidth]{figs/blob-diff-lumpy.pdf}

  % \includegraphics[width=0.3\textwidth]{figs/blob-unc-likelihood.pdf}\hfill
  % \includegraphics[width=0.3\textwidth]{figs/blob-unc-harmonic.pdf}\hfill\hspace{0.3\textwidth}

  % \includegraphics[width=0.3\textwidth]{figs/blob-ratio-likelihood.pdf}\hfill
  % \includegraphics[width=0.3\textwidth]{figs/blob-ratio-harmonic.pdf}\hfill\hspace{0.3\textwidth}

@@

We expect $T_\text{cad}$ to be a function of two dynamical time scales of the system: the rotational period of the asteroid $P_\omega$ and the time spent near perigee $T_p$. The latter can be estimated by
\begin{equation}
  T_p \sim \frac{r_p}{v_\infty}\brackets{2\frac{\mu_\mathcal{B}}{r_pv_\infty^2}+1}^{-\frac{1}{2}}
  \label{eqn:tp}
\end{equation}
which is the ratio of the perigee radius to velocity at perigee. The exact choice of the formula of $T_p$ not obvious, and alternatives to equation \ref{eqn:tp} are possible. For the simulated asteroid, $P_\omega = 9$ hr and $T_p = 42$ min by this definition.

%We further study the effect of $P_\omega$ and $T_p$ on $T_\text{cad}$ in appendix \ref{app:cadence-tests}, and find that both affect $T_\text{cad}$ by roughly the same amount.

>>>>

Over the past twenty years, the increase in quantity and quality of sensitive all-sky surveys has prompted the discovery of numerous asteroids. Such advances have been made via ground-based surveys such as the Catalina Sky Survey \cite{larson1998catalina}, Pan-STARRS \cite{kaiser2002pan}, and the Lincoln Near-Earth Asteroid Research project (LINEAR) \cite{stokes2000lincoln}, as well as space-based instruments such as the Wide-field Infrared Survey Explorer (WISE) mission \cite{wright2010wide}. Many of these asteroids are relatively small, but some are kilometre-sized and a few are predicted to closely encounter Earth or other planets in the near future. More encounter candidates are likely to be discovered by new efforts such as the Large-aperture Synoptic Survey Telescope (LSST) \cite{tyson2002large}. Their encounters can then be monitored by global ground-based networks such as the Las Cumbres Observatory (LCO) \cite{brown2013cumbres}. Such ground-based monitoring is typically used to derive the rotation period of an asteroid and its surface properties (see e.g. \cite{10.1093/mnras/stab1252}).

@@

\begin{figure*}
  \centering
  \includegraphics[width=0.33\textwidth]{figs/asym-ell-likelihood.pdf}\hfill
  \includegraphics[width=0.33\textwidth]{figs/asym-ell-harmonic.pdf}\hfill
  \includegraphics[width=0.33\textwidth]{figs/asym-ell-lumpy.pdf}

  \includegraphics[width=0.33\textwidth]{figs/asym-ell-diff-likelihood.pdf}\hfill
  \includegraphics[width=0.33\textwidth]{figs/asym-ell-diff-harmonic.pdf}\hfill
  \includegraphics[width=0.33\textwidth]{figs/asym-ell-diff-lumpy.pdf}

  \includegraphics[width=0.33\textwidth]{figs/asym-ell-unc-likelihood.pdf}\hfill
  \includegraphics[width=0.33\textwidth]{figs/asym-ell-unc-harmonic.pdf}\hfill\hspace{0.33\textwidth}

  \includegraphics[width=0.33\textwidth]{figs/asym-ell-rat-likelihood.pdf}\hfill
  \includegraphics[width=0.33\textwidth]{figs/asym-ell-rat-harmonic.pdf}\hfill\hspace{0.33\textwidth}

@@

\begin{figure*}
  \centering
  \includegraphics[width=0.33\textwidth]{figs/blob-likelihood.pdf}\hfill
  \includegraphics[width=0.33\textwidth]{figs/blob-harmonic.pdf}\hfill
  \includegraphics[width=0.33\textwidth]{figs/blob-lumpy.pdf}

  \includegraphics[width=0.33\textwidth]{figs/blob-diff-likelihood.pdf}\hfill
  \includegraphics[width=0.33\textwidth]{figs/blob-diff-harmonic.pdf}\hfill
  \includegraphics[width=0.33\textwidth]{figs/blob-diff-lumpy.pdf}

  \includegraphics[width=0.33\textwidth]{figs/blob-unc-likelihood.pdf}\hfill
  \includegraphics[width=0.33\textwidth]{figs/blob-unc-harmonic.pdf}\hfill\hspace{0.33\textwidth}

  \includegraphics[width=0.33\textwidth]{figs/blob-rat-likelihood.pdf}\hfill
  \includegraphics[width=0.33\textwidth]{figs/blob-rat-harmonic.pdf}\hfill\hspace{0.33\textwidth}

@@

We expect $T_\text{cad}$ to be a function of two dynamical time scales of the system: the rotational period of the asteroid $P_\omega$ and the dynamical time scale of the orbit (the latter can be estimated in multiple ways, since both $r_p / v_\infty$ and $v_\infty r_p^2 / \mu_\mathcal{B}$ have units of time and may be relevant). How these time scales affect $T_\text{cad}$ is discussed in appendix \ref{app:cadence-tests}.
% \begin{equation}
%   T_p \sim \frac{r_p}{v_\infty}\brackets{2\frac{\mu_\mathcal{B}}{r_pv_\infty^2}+1}^{-\frac{1}{2}}
%   \label{eqn:tp}
% \end{equation}
% which is the ratio of the perigee radius to velocity at perigee. The exact choice of the formula of $T_p$ not obvious, and alternatives to equation \ref{eqn:tp} are possible. For the simulated asteroid, $P_\omega = 9$ hr and $T_p = 42$ min by this definition.
